\clearpage
\thispagestyle{empty}
Jens Christensen

List of Corrections

Overall Corrections

\begin{enumerate}
  \item All notes in thesis corrected.
  \item Labels added to figures missing it and colours changed in Figures~\ref{fig:SeparationWaveformEx320.pdf} (pp. \pageref{fig:SeparationWaveformEx320.pdf}), \ref{fig:SeparationWaveformEx640.pdf} (pp. \pageref{fig:SeparationWaveformEx640.pdf}), \ref{fig:SeparationWaveformExBig320.pdf} (pp. \pageref{fig:SeparationWaveformExBig320.pdf}) and \ref{fig:SeparationWaveformExBig640.pdf} (pp. \pageref{fig:SeparationWaveformExBig640.pdf}).
  \item /todo
  \item Done. See beginning of thesis.
  \item
\end{enumerate}

Abstract

Chapter 2
\begin{enumerate}
\item \gls{mfcc} has been added in the literature review (pp. \pageref{corrections:mfcc1}) as a major audio feature in speech feature extraction and as proposed future work (pp. \pageref{corrections:mfcc2}).
    \item On pp. \pageref{corrections:viterbi} the Viterbi algorithm is introduced as a recursive optimal solution to the problem of estimating the state sequence of a discrete-time finite-state Markov process. A sentence has been added to clarify how the Viterbi may be used for a classification application given certain circumstances.
    \item On pp. \pageref{fig:LitRev_DetectCompare} and \pageref{fig:LitRev_DetectCompare2} references have been added to indicate the inspiration for the particular method's use in the comparison.
    \item The patent mentioned in the list of corrections was already referenced (a long with other occurrences of the same IP) in reference \cite{Seltzer2011Patent} and does not appear to be related to patent \cite{US8233353}.
        
    \end{enumerate}
